Automatically generated NE tagged corpora for English and Hungarian

### 2012-03-25 paper/engNERwiki.tex
jav

\begin{description}
\item[Derived words] According to the CoNLL guidelines, word derived from name entities must be tagged as \texttt{MISC}. We complied with this rule by tagging each entity whose head is not a noun, as well as when the link's anchor text is not contained in the entity's name, as \texttt{MISC}. The most prominent example for such entities are ethnic groups, which can be linked to their home country, a \texttt{LOC}; e.g. \textit{Turkish} to \textit{Turkey}. Our solution assigns the correct tag to these entities.
\item[First word in a sentence] As first words are always capitalized, labelling them is difficult if they are unlinked and not contained in the entity alias set. We base the decision on the POS tag of the first word: if it is \texttt{NNP}, we tag it as \texttt{UNK}; otherwise, \texttt{O}.
\item[Reference cleansing] Page titles and anchor texts may contain more than just the entity name. Personal titles are part of the entity name in Wikipedia, but not in CoNLL, and punctuation marks around the entity may become part of the link by mistake. We tag all punctuation marks after the entity name as \texttt{O}.

@@

Certain types of sentences were filtered from the corpus. As mentioned above, sentences with words tagged as \texttt{UNK} are discarded. Furthermore, there are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc. On the one hand, these sentence fragments may prove to be of too low quality to be of any use in the traditional NER task. On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is known to be noisy and fragmented. As a compromise solution, we included these fragments in the corpus, but labelled them as "low quality", so that users of the corpus can decide whether they want to use them or not. A sentence is labelled as such if it either lacks a punctuation mark at the end, or it contains no finite verbs.

@@

To annotate our corpus with NE-tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, Szeged NER corpus \cite{Szarvas:06}. It is quite similar to the CoNLL standards: it contains business newswire texts, it comprises ca.~200,000 tokens, and the text has been annotated with NE class labels in line with the CoNLL annotation scheme. However, the convention of what constitutes a NE is slightly different for Hungarian.

@@

Having the obvious advantages, an automatically generated corpus can not serve as a gold standard dataset. Then what can we do with these so called "silver standard corpora" (SSC)? They can be very useful for improving NER in more ways: (a) for less resourced languages, they can serve as training corpora in absence of gold standard datasets; (b) they can serve as (supplementary or independent) training sets for domains differing from newswire; (c) they can be sources of huge entity lists, and (d) feature extraction.

@@

We evaluated our English corpus against (a) itself (divided into train and test sets); and (b) a manually annotated English corpus. Since the filtered English Wikipedia corpus, containing only the sentences with NEs, is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus). It was divided into train and test sets (90\%-10\%).

For the English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format. It is well-studied that cross-checking between different corpora decrease F-measure. Stylistic and domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems. Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs much better on corresponding test sets than on test sets from other sources. The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see Table \ref{enresults} for results).

@@

Since in the subtask of NE labeling for Hungarian we followed the Szeged NER corpus annotation guidelines, it is obvious to perform the experiments on this dataset, as well. Hungarian results are quite similar to the English ones (see Table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher, than ones for English. It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency and diversity of this class is smaller (see section \ref{sec:hun}). In contrast to the CoNLL corpus, the Szeged NER corpus was highly accurately annotated with an inter-annotator agreement over 99\%.

@@

We have presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from Wikipedia. In contrast to the methods used so far for automatic annotation of NEs in Wikipedia texts, we applied a new approach, namely mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to Wikipedia entities. Following Nothman \shortcite{Nothman:08}, the process can be divided into four main steps: (1) classifying all articles into entity classes; (2) splitting Wikipedia articles into sentences; (3) labeling NEs according to link targets; and (4) selecting sentences for inclusion in the corpus.

@@

Automatically generated, so called "silver standard corpora" can be very useful for improving NER in more ways. We showed that huge gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves overall F-measure. Moreover, our Hungarian corpus can serve as a training corpus to build NE taggers for more general domains, than the classic newswire.

>>>>

\begin{description}
\item[Derived words] According to the CoNLL guidelines, word derived from NEs are tagged as \texttt{MISC}. We complied with this rule by tagging each entity whose head is not a noun, as well as when the link's anchor text is not contained in the entity's name, as \texttt{MISC}. The most prominent example for such entities are nationalities, which can be linked to their home country, a \texttt{LOC}; e.g.~\textit{Turkish} to \textit{Turkey}. Our solution assigns the correct tag to these entities.
\item[First word in a sentence] As first words are always capitalized, labeling them is difficult if they are unlinked and not contained in the entity alias set. We base the decision on the POS tag of the first word: if it is NNP, we tag it as UNK; otherwise, O.
\item[Reference cleansing] Page titles and anchor texts may contain more than just the entity name. Personal titles are part of the entity name in Wikipedia, but not in CoNLL, and punctuation marks around the entity may become part of the link by mistake. We tag all punctuation marks after the entity name as \texttt{O}.

@@

As mentioned above, sentences with words tagged as \texttt{UNK} are discarded. Furthermore, there are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc. On the one hand, these sentence fragments may prove to be of too low quality to be of any use in the traditional NER task. On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is known to be noisy and fragmented. As a compromise solution, we included these fragments in the corpus, but labelled them as "low quality", so that users of the corpus can decide whether they want to use them or not. A sentence is labelled as such if it either lacks a punctuation mark at the end, or it contains no finite verbs.

@@

To annotate our corpus with NE tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, Szeged NER corpus \cite{Szarvas:06}. It is similar to CoNLL standards: contains newswire texts, comprises ca.~200,000 tokens, and is annotated with NE class labels in line with the CoNLL annotation scheme. However, the convention of what constitutes a NE is slightly different for Hungarian.

@@

Having the obvious advantages, an automatically generated corpus can not serve as a gold standard dataset. Then what can we do with silver standard corpora? They can be very useful for improving NER in more ways: (a) for less resourced languages, they can serve as training corpora in absence of gold standard datasets; (b) they can serve as (supplementary or independent) training sets for domains differing from newswire; (c) they can be sources of huge entity lists, and (d) feature extraction.

@@

The English train corpus was evaluated against itself and a manually annotated English corpus. Since the filtered English Wikipedia corpus, containing only the sentences with NEs, is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus). It was divided into train and test sets (90\%-10\%).

For English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format. It is well-studied that cross-checking between different corpora decrease F-measure. Stylistic and domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems. Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs much better on corresponding test sets than on test sets from other sources. The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see Table \ref{enresults} for results).

@@

Since in NE labeling for Hungarian we followed the Szeged NER corpus annotation guidelines, it is obvious to perform the experiments on this dataset, as well. Hungarian results are quite similar to the English ones (see Table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher, than ones for English. It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency and diversity of this class is smaller (see section \ref{sec:hun}). In contrast to the CoNLL corpus, the Szeged NER corpus was highly accurately annotated with an inter-annotator agreement over 99\%.

@@

We have presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from Wikipedia. In contrast to the methods used so far for automatic annotation of NEs in Wikipedia texts, we applied a new approach, namely mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to Wikipedia entities. Following Nothman \shortcite{Nothman:08}, the process can be divided into four main steps:  classifying all articles into entity classes; splitting Wikipedia articles into sentences; labeling NEs according to link targets; and selecting sentences for inclusion in the corpus.

@@

Automatically generated, silver standard corpora can be very useful for improving NER in more ways. We showed that huge gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves overall F-measure. Moreover, our Hungarian corpus can serve as a training corpus to build NE taggers for more general domains, than the classic newswire.
